---
title: "Lab Week 5 time series and a map"
author: "Me"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)
```

## Monthly US energy consumption (renewables)

```{r}
us_renew <- read_csv(here("data","renewables_cons_prod.csv")) %>% 
  janitor::clean_names()
#unique function to see what each unique variable is in a column 
```

- Make description all lower case
- Only keep observations where the "description" variable only contains "consumption"
- Remove any variables where "description contains "total"

```{r}
renew_clean <- us_renew %>%
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total")) #! means do the opposite

```

### Convert 'yyyymm' column to a date
```{r}
renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

#automatically add 01 as the day, but we will remove 
#tsibble created cool year/month column for it now 

#make a version where month and year are in seperate columns
renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))
```

#Now that it is cleaned up, lets go ahead and look at it 

```{r}
renew_gg <- ggplot(renew_date, aes(x=month_sep, 
                                   y=value,
                                   group = description)) +
  geom_line(aes(color = description))

renew_gg

#view(palettes_d_names) to see different color themes
```
Updating colors with paletteer palettes:
```{r}
renew_gg +
  scale_color_paletteer_d("nationalparkcolors::CraterLake")
```
### Coerce renew_parsed to a tsibble

```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, 
                       index = month_sep)
#letting the package know that the tsibble package is running month_sep as the time frame
```

Let's look at our data in a couple of different ways:
```{r}
renew_ts %>% autoplot(value) #autoplot is a good one for tsibble
renew_ts %>% gg_subseries(value) #year by month by energy
renew_ts %>% gg_season(value) #just a time saver, worked for me, but we will also make it in ggplot 
ggplot(data = renew_parsed, aes(x=month, y=value, group=year)) +
  geom_line(aes(color = year)) +
  facet_wrap(~description, 
             ncol = 1,
             scales = "free",
             strip.position = "right")


```

### Lets just look at hydroelectric energy consumption 

```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)
hydro_ts %>% gg_season(value)

#ugly version of ggseason 
ggplot(hydro_ts, aes(x=month, y=value, group=year))+
  geom_line(aes(color = year))
```

### What if I want the quarterly average consumption for hydro?

```{r}
#not group by but index by, just more compatible with tsibble 

hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% 
  summarize(avg_consumption = mean(value)) #now gives the data by quarter, pretty cool 
```

#Decompose that hydro_ts

```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~season(window=5))) 

components(dcmp) %>% autoplot()
```

```{r}
hist(components(dcmp)$remainder)
```

Now lets loo at ACF
```{r}
hydro_ts %>% 
  ACF(value) %>% 
  autoplot()
```

#Forecast DANGER DANGER DANER

```{r}
hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value)
  ) %>% 
  fabletools::forecast(h = "4 years")

hydro_model %>% autoplot()

hydro_model %>% 
  autoplot(filter(hydro_ts, year(month_sep) > 2010))
```

### Make a world map! 

```{r}
world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)
```

